---
title: "Probability of Shared Birthdays"
author: "Patrick Kelly"
date: "July 26, 2015"
output:
  pdf_document:
    toc: yes
---

# INTRODUCTION
## Question: Given 30 randomly selected singleton (no twins) people, what is the probability that at least 2 of them have the same birthday? The answer might surprise you. Salman Khan finds the likelihood of this occuring, using permutations and a virtual TI - 85 calculator.
[KhanAcademy](http://tinyurl.com/SharedBirthdays)

[Click here](https://www.khanacademy.org/math/precalculus/prob-comb/prob-combinatorics-precalc/v/birthday-probability-problem)"English"


### Wikipedia also explores the problem and possible solutions.
[Wikipedia]http://tinyurl.com/BirthdayMatches

###P(At least 2 people share a birthday) = 1 - (365! / (365-30)! / 365^30^) = 0.7063
##Since R has a problem with huge factorials, this program calculates the expected mean probability and its 95% confidence interval, using another approach.
***
### Choose the numbers for people, repeats and iterations.
```{r}
people <- 20
repeats <- 1000
iterations <- 200
```
##You chose `r people` people, `r repeats` repeats and `r iterations` iterations.
###Create a vector of numbers 1 through 365 and 3 empty dataframes.
<div class="pagebreak"></div>
```{r}
days <- matrix(c(1:365))
prob1 <- data.frame()
prob.at.least.1 <- data.frame()
prob2 <- data.frame()
```
###Create a function that finds the average probability of shared birthdays for a given number of people and repeated trials.

```{r}
SameBirthday <- function (people, repeats) {
  for (i in 1:repeats){
    sampl <- days[sample(days, people, replace = TRUE),]
    diff.birthday <- length(unique(sampl)) < people
    prob1 <- rbind(prob1,diff.birthday)}
  trues <- sum(prob1$TRUE. == TRUE) + sum(prob1$FALSE. == TRUE)
  trials <- repeats
  options(digits=3)
  prob.at.least.1 <- trues / trials * 100
  prob.at.least.1 <<- data.frame(prob.at.least.1) 
  # <<- variable is available outside of the function
  # cat("Probability of at least 1 match =",
  # prob.at.least.1,"%","(",trials,"trials )", "\n")
}
```
###Calculate the estimated average probability for `r people` people in `r repeats` trials.
```{r}
SameBirthday(people,repeats)
```
## Probability of at least 1 match = `r prob.at.least.1`% (`r repeats` trials)
###Now create another function that iterates the first one `r iterations` times to calculate an estimate of an overall average mean statistic of shared birthdays for `r people` people, and the 95% confidence interval.
```{r}
OverallMean <- function(iterations) {
  for (i in 1:iterations){
    SameBirthday(people,repeats) 
    prob2 <<- rbind(prob2,prob.at.least.1)}
    options(digits=3)
  mean.prob <<- mean(prob2$prob.at.least.1)
  sd.prob <- sd(prob2$prob.at.least.1)
  sem.prob <- sd.prob / sqrt(iterations)  # Standard error of the mean
  t <- qt(0.975,df = iterations-1)
  marg.error <- t * sem.prob  # Margin of error of the mean
  ci.prob <<- mean.prob + c(-marg.error, marg.error)
  cat(" Mean probability of at least 1 match = ",mean.prob,"\n")
  cat(" 95% confidence interval for the mean:","\n",  
    "Lower Limit =", ci.prob[1],"      Upper Limit = ",ci.prob[2])
} 
```
###Run the second function for `r iterations` iterations.
```{r}
OverallMean(iterations)
```   
##Salman Khan's estimate for the likelihood of shared birthdays for 30 random people is 70.6%. Is this estimate included in the 95% confidence interval?
***
##Visualization of the results
###First a boxplot, which may or may not show outliers (circles), defined by John Tukey as values more extreme than 1.5 times the Inner Quartile Range
```{r}
boxplot(prob2$prob.at.least.1)
```

###Finally a histogram of the frequencies with a vertical line indicating the overall mean and an overlying density plot.
```{r}
hist(prob2$prob.at.least.1, 
  col = "grey", prob = T,
  main = "Distribution of Estimated Means", 
  xlab = "% Probability of at least one match", 
  sub ="Blue Line = Overall Mean")
abline(v = mean.prob, col = "blue",lwd=2)
abline(v = ci.prob[1], col = "green",lwd=2)
abline(v = ci.prob[2], col = "green",lwd=2)
lines(density(prob2$prob.at.least.1), col="red", lwd=2) 
```   

